
---
title: "Guppy Network Infection Experiment"
author:
- David R. Clark
- Jason Walsman
- Faith Rovenolt
- Jessica Stephenson
date: "01/31/2023"
output:
  pdf_document:
    toc: yes
    number_sections: yes
  editor_options:
    chunk_output_type: console
  word_document:
    toc: yes
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)
```

\section{Introduction}

\subsection{Overall summary}

This is an analysis to examine how host traits, particularly host sociality and sex, impacts the spread of disease within experimental epidemics of Trinidadian guppies and Monogenean parasites. 

\subsection{Quick experimental summary}

We released 6 female guppies and 3 male guppies into experimental behavioral enclosures together and recorded their baseline sociality (Here defined as Contact Rate) for 2 days. We then introduced an infected female index fish on day 3 to see how quickly parasites spread within the group of fish and how likely specific individuals to pick up parasite infection. We repeated recording behavior for 7 days post infection for a total of 9 days of behavioral recordings, however the analysis within this document is limited to days 1-3. 

\subsection{Questions}

\subsubsection{Individual level}

- Are females more likely to be infected than male hosts?

- Does host worm contact rate explain individual level probability of infection?

- Do infected index fish receive less contact rate from uninfected conspecifics relative to uninfected index fish?

-Do infected index fish with higher infection intensity receive less contact rate from uninfected conspecifics?

\subsection{Data description, structure and type}

This analysis uses a single master dataframe titled:DyadindividualContacts_20240726.csv that is subset into different portions based on what questions/analysis we are addressing. Variables within the data frame as follows:

\textbf{Ind}: The individual ID of each fish used in the trial. \newline
 \textbf{ContactRate}: The daily average contact rate for each individual  ($contacts/dyad/min$) \newline
 \textbf{Day3ContactRatewoI}: The daily average contact rate for each non-index individual with other non-index individuals. Index fish is the fish introduced on Day 3 with infection (or not for uninfected conrols) ($contacts/dyad/min$) \newline
 \textbf{IndexContact}: The daily average contact rate with the index individual. Repeated for each individual since they only had one day with the index individual. ($contacts/dyad/min$) \newline
 \textbf{Day}: The day in which the behavior was taken. Days 1 and 2 are days without the infected index individual and day three is with the infected index individual. 3 levels. One level for each day. \newline
 \textbf{Popavg}: The average contact rate for each group ($Contacts/dyads/min/fish$) \newline
 \textbf{Pop}: The group ID for each replicate. One level per group. \newline
\textbf{Treatment}: Whether the group received an infected or uninfected index fish at the beginning of day 3. 2 levels. \newline
\textbf{TempTreatment}: The treatment (infected or uninfected) of each group for each day. For every group it will be Unf on days 1 and 2 due to not having an index, but Inf for Infected treatments for days 3. 2 levels. \newline
\textbf{Sex}: The sex of each individual. 2 levels, F for female, M for male. \newline
\textbf{Inf.}: This is renamed to Infection in some subset data sets. Whether or not an individual was infected at the end of day 3. 2 levels, 1 - infected, 0-uninfected \newline
\textbf{worms}: The number of worms on each individual at the end of day 3. \newline
\textbf{prev}: Prevalence (proportion of infected individuals) for the group at the end of day 3. \newline
\textbf{IndexWorm}: Number of worms on the index fish at the end of day 3. This is used as a metric for the amount of worms an individual will have been exposed to during intitial epidemic spread.  \newline
\textbf{Index}: Whether the individual was an index fish or not. 2 levels, 0-not index, 1-index  \newline
\textbf{ChConWithIn}: Change in contact with including the contacts with the index between days 2 and 3. \newline
\textbf{ChConWOIn}: Change in contact with with only non-index individuals between days 2 and 3. \newline
\textbf{ChConPop}: Change in the group average contact rate between days 2 and 3 \newline
\textbf{PreSVL}: The snout to vent length of the guppy prior to introduction into behavioral arena. Measured in ($Millimeters$) \newline
\textbf{PreWeight}: The weight of the individual prior to introduction into the behavior box. Measured in ($grams$) \newline
\textbf{PostSVL}: The snout to vent length of the guppy after wrapping up entire experiment.  Measured in ($Millimeters$)  \newline
\textbf{PostWeight}: The Weight of the guppy after wrapping up entire experiment.  Measured in ($grams$) \newline



```{r importing data and defining key metrics and loading libraries, warning=F,message=FALSE}
#Load in libraries for analysis and data wrangling
#Visualization
library(ggplot2)
library(visreg)
source("http://highstat.com/Books/BGS/GAMM/RCodeP2/HighstatLibV6.R")
#(generalized) Linear mixed modeling
library(lme4)
library(glmmTMB)
#Statistcal analysis reporting and model validation
library(performance)
library(car)
library(lmtest)
library(DHARMa)
#Data wrangling
library(dplyr)
library(plyr)
library(tidyverse)
library(emmeans)


#Importing the master data sheet check intro for description of factors
#Choose X datasheet (specify when finished)
IndContacts <- read.csv("DyadindividualContacts_20240726.csv")

#This is creating a variable to calcualte the contact rate with the infected individual for non-infected fish and combining them back together. 
#Subsetting to inidivudals who are not the index
#Addtitionally this code subsets down to only focusing on dyads with the index fish as the recipient

IndContactsNI <- IndContacts %>%
  filter(Index != "1" & day == "3" & str_detect(recipID2, "_10"))%>%
  dplyr::rename(IndexContactR = ContactInitR)%>%
  select(c(fishID2, recipID2, IndexContactR))

#Subseting to individuals who are the index fish
IndContactsI<- IndContacts %>%
  filter(Index == "1")%>%
  mutate(IndexContactR = NA )%>%
  select(c(fishID2, recipID2, IndexContactR))
#Rbinding our two variables together
IndContactsNI<-rbind(IndContactsNI, IndContactsI)

#Merging our dataframe back together by fishID to add in the indexcontactrate
IndContacts2<-merge(IndContacts, IndContactsNI, by = c("fishID2"), .keep_all=TRUE)%>%
  #select(-c(recipID2.y, day.y, population.y))%>%
  dplyr::rename(recipID2 = recipID2.x)%>%
  distinct(fishID2, recipID2, population, day,.keep_all=TRUE)

#Sum contact rate across all videos in a day for each dyad
#We want to combine all contacts across a day. We added 0.001 to each of these to help with further model fit without impacting the distribution of the data
IndContactsSum <- IndContacts2%>%
  select(-c(X, fishID, frame_num, recipID))%>%
  mutate(ContactInitR = ContactInitR + 0.0001,
         TotalCR = TotalCR + 0.0001,
         RecipCR = RecipCR + 0.0001)%>%
  group_by(fishID2, recipID2, day)%>%
  dplyr::summarise(ContactInitRs = sum(ContactInitR), 
            TotalCRs = sum(TotalCR), 
            RecipCRs = sum(RecipCR),
            conttactinit = sum(contactinit),
            RecipfledCR = sum(RecipfledCR))

#Remove unneeded columns
IndContactssub<-IndContacts2%>%
  select(-c(X, fishID, frame_num, recipID, ContactInitR, TotalCR, RecipCR, RecipfledCR))%>%
  group_by(fishID2, day)

#Merge sheets together so that new data frame has all variables
IndContacts3<-merge(IndContactsSum, IndContactssub, by = c("fishID2", "recipID2",  "day"), .keep_all=TRUE)

#Going down to one row per indiviual
IndContacts3 <- IndContacts3 %>%
  group_by(fishID2,recipID2,  population, day)%>%
  distinct(fishID2,recipID2,  population, day, .keep_all = TRUE)%>%
  filter(fishID2 != "RCinf1_NA")

#Calculating important metrics that may apply across all subsetting of data

#Wormcontacts = Number of contacts a fish has had with infectious agents. So the multiplicative of contact rate with the infected index with the number of worms on the infected index
IndContacts3 <- IndContacts3 %>%
  mutate(
    WormContact = IndexContactR * IndexWorm
    )
#Assigning a fish label to the fish so that if we want to subset down to one metric for fish because not all metrics change across days. This currently doesnt work due to index only have day 3 measure
#Fish<-c("1","0","0")

#Adding this factor to overall data frame
#PopulationContacts$Fish<-as.factor(Fish)

#Overall summary just checking that everything looks okay.
summary(IndContacts3)

#Set data strcuture for some factors

#Setting Day as a factor instead of a numeric
IndContacts3$day <- as.factor(IndContacts3$day)
#Setting whether a fish was an index (intitial infection) or not as factor
IndContacts3$Index<-as.factor(IndContacts3$Index)
#Setting whether an fish was infected or not as factor 
IndContacts3$Infectionstat <- as.factor(IndContacts3$Infectionstat)
#Setting whether fish were male or female as a factor
IndContacts3$Sex <- as.factor(IndContacts3$Sex)
IndContacts3$recipsex <- as.factor(IndContacts3$recipsex)
#Setting the treatment of fish group (infected index or not) as factor
IndContacts3$InfectionTrt <- as.factor(IndContacts3$InfectionTrt)
#Setting population  the fish were from as a factor
IndContacts3$population<-as.factor(IndContacts3$population)
#Setting fish ID as a factor
IndContacts3$fishID2<-as.factor(IndContacts3$fishID2)
#Setting recipient ID as a factor
IndContacts3$recipID2<-as.factor(IndContacts3$recipID2)



```


\subsection{Initial data exploration}

```{r initial data exploration, warning=F}
#Pairs plot to look at autocorrelation and potential relationships before diving into analysis
pairs(~TotalCRs+ContactInitRs+IndexContactR+Sex+Infectionstat+IndexWorm+InfectionTrt, lower.panel=panel.smooth, diag.panel=panel.hist, 
upper.panel=panel.cor, data=IndContacts3)



#Checking to see if there are any day differences
daylm<-glmmTMB(TotalCRs~day+(1|population), family=beta_family(), IndContacts3)
#summary of the model
summary(daylm)
#Testing for significance
drop1(daylm,test="Chisq")
#It seems we have some high contact rate on day 1, we have to eliminate it due to acclimation


#Ploting behavior by individual across groups
daylmVR<-visreg(daylm, "day", scale="response", partial=TRUE, plot=FALSE)
#Generate the fitted relationship
daylmVRfit<-daylmVR$fit
#Generate the residuals
daylmVRres<-daylmVR$res
#Plot this relationship
ggplot(daylmVRres, aes(x=day,y=visregRes))+geom_boxplot(outliers=FALSE)+geom_jitter(alpha=.2)+theme_classic()+theme(legend.position="none", text=element_text(size=22))+ylab("Total contact rate")+xlab("Day")


#There are significant differences in day so we have to subset out first day
#We need to remove first day measurements. 
IndContactPW<-IndContacts3%>%
filter(day == "2" | day == "3" & Index == "0" & recipID2 != str_detect(recipID2, "_10"))%>%
select(fishID2, day, ContactInitRs)%>%
  pivot_wider(names_from = day, values_from = ContactInitRs)%>%
  na.omit(2)%>%
  mutate(ChContactinit = `3` - `2`)%>%
  select(fishID2, recipID2, ChContactinit)
#Create a index only data frame and set its change in contact rate as NA given it only came in on day 3
IndContactI<-IndContacts3%>%
  filter(Index == "1")%>%
  mutate(ChContactinit = NA)

  
IndContacts4 <- merge(IndContactPW, IndContacts3, by=c("fishID2", "recipID2", "population"))%>%
  filter(day != "1")

IndContacts4<-rbind(IndContacts4, IndContactI)


# %>%
# filter(population == "ADinf1" | population == "ADunf1" | population == "GZinf1" | population ==  "GZunf1"| population == "SBinf1"| population =="SBunf1")
```

\section{Individual Level Analysis}



\subsection{Are hosts with higher worm contact rate more likely to become infected?}

For this first bit we need to select the data we need to answer this specific question
```{r selection and subseting to data needed for this analysis, warning=F}
#Filtering down to only infected treatments, individuals who are not the index, one measure per fish and days 2 and 3. Note the days 2 and 3 is because intitial viewing of behaviors across days looks like fish were not behaving normally on day 1 due to acclimation to the new environment.
IndContactsPI1<-IndContacts3 %>%
  filter(Index != "1" & day == "3" & InfectionTrt == "1" & Sex =="F" & str_detect(recipID2, "_10"))

```


\subsubsection{Fitting the contact rate versus probability of infection model}

We're fitting a binomial generalized linear mixed model with whether an individual was infected or not at the end of day 3 as our response variable and worm contact rate as predictor variables. We have population as a random effect to control for non-independence of worm contact rate between individuals within the same group.

In addition to this, we fit a second GLMM examining the same relationship but with host contact rate instead of the product term of worm contact rate to see if it better fits our data.

```{r fitting the glmmTMB for prob of infection and contact rate, warnging=F}
#generalized linear mixed model to determine if individual sociality explained the probability of becoming infected
ProbinfLMF<-glmmTMB(Infectionstat ~ 
                      WormContact+
                      I(WormContact^2)+
                      (1|population),
                    family=binomial(), 
                    data=IndContactsPI1)


#We fit a second model with just contact rate to see if it does a better job than fitting the product term of contact rate and infection intensity of the index
#A model with just social contacts to test against using the social contacts 
ProbinfLMFSC<-glmmTMB(Infectionstat ~ 
                      ContactInitRs+
                      I(ContactInitRs^2)+
                      (1|population),
                    family=binomial(), 
                    data=IndContactsPI1)

#Summary of our two models
summary(ProbinfLMF)
summary(ProbinfLMFSC)

```

\subsubsection{Next we see will validate the model to make sure everything is looking good.}

```{r model validation for prob of infection vs ContactRate, warning=F, plot=F}
#Using DHARMa package to look at quantile residuals to validate the worm contact rate model fit
sim_residuals_ProbinfLM <-simulateResiduals(ProbinfLMF, 1000) 
#Quantile residuals
#Ploting the quantile residuals
plot(sim_residuals_ProbinfLM) 
#Testing dispersion
testDispersion(sim_residuals_ProbinfLM)

#Using DHARMa package to look at quantile residuals to validate the contact rate model fit
sim_residuals_ProbinfLMSC <-simulateResiduals(ProbinfLMFSC, 1000) 
#Quantile residuals
#Ploting the quantile residuals
plot(sim_residuals_ProbinfLMSC) 
#Testing dispersion
testDispersion(sim_residuals_ProbinfLMSC)

#Below we are exmaining the quantile residuals among the fixed variables to see if there are any funky patterns. Excluded from main run for space but remove # to see plots
#plotResiduals(sim_residuals_InfCRlm, IndividualInfDFD2$Sex)
#plotResiduals(sim_residuals_InfCRlm, IndividualInfDFD2$ContactRate)

check_model(ProbinfLMF)
```

\subsubsection{Likelihood ratio test to test for significance in our model}

```{r Likelihood ratio test for prob of inf with Sociality, warning=F}

#Drop 1 with chisq test to do a likelihood ratio test for our model
drop1(ProbinfLMF, test="Chisq")

```

\subsubsection{Visualization of the relationships between worm contact rate and probability of infection}

```{r Visualization of the relation for prob inf and contactrate/sex, warning=F}
#Visreg of contact rate effect
InfWCRvr<-visreg(ProbinfLMF, "WormContact", scale="response", partial=T, overlay=TRUE, gg=TRUE)+xlab("Worm contact rate (worms x contact frames/total frames) ")+ylab("Chance of being infected on day 3")+theme_classic()+theme(text=element_text(size=18))
print(InfWCRvr)

```

\subsection{Do hosts contact uninfected index as much as infected index individuals?}

```{r selection and subseting to data needed for index worm and contact analysis, warning=F}
#Filtering down to only 
#Getting index contact rate
IndContactsIndexrecip <- IndContacts4 %>%
  filter(Index == "1" )
  
```
\subsubsection{Fitting a model to test for total contact rate for infected versus uninfected index individuals}

```{r fitting the model for infected and uninfected index}
#Logging and scaling contact rate due to heavy skew and the desire to be comparable between explanatory variables
IndContactsIndexrecip$ScContactInitRs<-scale(log10(IndContactsIndexrecip$ContactInitRs))

#Fitting the glmmTMNB for received contacts by infection treatment of the group
IndexTClm<-glmmTMB(RecipCRs~
                     InfectionTrt+
                     recipsex+
                     ScContactInitRs+
                     I(ScContactInitRs^2)+
                     (1|fishID2),
                   family=beta_family(link="logit"), 
                   data=IndContactsIndexrecip)


#summary of the model
summary(IndexTClm)

```

\subsubsection{Validating the recipient initiated contact with index model}
```{r validating the model fit for the totalcontact overall}

sim_residuals_IndexTClm <-simulateResiduals(IndexTClm, 1000)  #Quantile residuals
#Ploting the quantile residuals
plot(sim_residuals_IndexTClm) 
#Testing dispersion
testDispersion(sim_residuals_IndexTClm)


#using the checkmodel function to validate model
check_model(IndexTClm)


# sim_residuals_IndexInitTClm <-simulateResiduals(IndexInitTClm, 1000)  #Quantile residuals
# #Ploting the quantile residuals
# plot(sim_residuals_IndexInitTClm) 
# #Testing dispersion
# testDispersion(sim_residuals_IndexInitTClm)

```

\subsubsection{Using a likelihood ratio test to test for significance in the recipient initiated contact with index model}

```{r testing for significance for the glmmTMB contact overall model}

drop1(IndexTClm, test="Chisq")


```


\subsubsection{Visualizing the relationship between recipient sex and total contacts with index}

```{r visualing the Infection Treatment variation relationship for the total contact glmmTMB model}
#Generating fitted relationship and residuals for contact rate and infection treatment
RCind<-visreg(IndexTClm, "ScContactInitRs", "InfectionTrt", scale="response", partial=T, plot=FALSE)

#Grabbing the fitted relationship
RCindfit<-RCind$fit
#Grabbing the partial residuals
RCinfres<-RCind$res
#Setting the color scheme for infected and uninfected groups
cpinf=c("darkgray", "purple")

#Plotting the relationship for received contact rate and infection treatment

ggplot(RCindfit, aes(x=ScContactInitRs, y=visregFit, group=InfectionTrt, colour=InfectionTrt, fill=InfectionTrt))+
  geom_smooth(data=RCindfit, method="loess", aes(x=ScContactInitRs, y=visregFit))+
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha=.3)+
  geom_point(data=IndContactsIndexrecip,aes(x=ScContactInitRs, y=RecipCRs))+
  theme_classic()+
  xlab("Scaled log initiated contact rate by index")+
  ylab("Initiated contact rate by non-index")+
  theme(text=element_text(size=22))+
  theme(legend.position="top")+
  scale_color_manual(values =cpinf )+
  scale_fill_manual(values =cpinf )


```

\subsection{Do index fish with higher infection intensity have less recipient contacts?}

Subsetting the data frame down for this analysis

```{r subsetting down to data for infection intensity and received contacts analysis}
#subsetting down
IndContactsIndexinfrecip <- IndContactsIndexrecip %>%
  filter(InfectionTrt == "1")

#Sclaing some of our variables for the analysis
#Index infection intensity
IndContactsIndexinfrecip$ScIndexWorm<-scale(IndContactsIndexinfrecip$IndexWorm)


```

\subsubsection{Fitting our GLM for the receipient contacts by infection intensity}
```{r fitting the linear model for infected index recipient contacts by worm load}

#Fitting the glmmtmb for index received contacts by infection intensity
IndexRCinflmrecip<-glmmTMB(RecipCRs~ScIndexWorm+
                             I(ScIndexWorm^2)+
                             recipsex+
                             ScContactInitRs+
                             I(ScContactInitRs^2)+
                             ScIndexWorm:ScContactInitRs+
                             ScIndexWorm:recipsex+
                             I(ScIndexWorm^2):recipsex, 
                           family=beta_family(link="logit"), 
                           data=IndContactsIndexinfrecip)

#summary of model
summary(IndexRCinflmrecip)

```

\subsubsection{Validating the recipient contact rate by worm load model}
```{r validating the model fit for the indexworm load and recipient contact model}

sim_residuals_IndexRCinflmrecip <-simulateResiduals(IndexRCinflmrecip, 1000)  #Quantile residuals

#Confirming there is no differences in median between groups
plot(sim_residuals_IndexRCinflmrecip$fittedResiduals~IndContactsIndexinfrecip$population )

#Ploting the quantile residuals
plot(sim_residuals_IndexRCinflmrecip) 
#Testing dispersion
testDispersion(sim_residuals_IndexRCinflmrecip)

# 
# sim_residuals_IndexRCinflminit <-simulateResiduals(IndexRCinflminit, 1000)  #Quantile residuals
# #Ploting the quantile residuals
# plot(sim_residuals_IndexRCinflminit) 
# #Testing dispersion
# testDispersion(sim_residuals_IndexRCinflminit)



```

\subsubsection{Using a likelihood ratio test to test for significance in recipient contact by index worm load model}
```{r testing for significance for the glmmTMB Index worm and contact model}

drop1(IndexRCinflmrecip, test="Chisq")

```


\subsubsection{Visualizing the relationship between recipient sex and recipient contact and worm load model}
```{r visualing the sex variation relationship for the total contact glmmTMB model}
#Extracting fit and residuals from our model
IndexRCinflmVR<-visreg(IndexRCinflmrecip,"ScIndexWorm", "recipsex", scale="response", partial=T, plot=FALSE)
#Extracting fit
IndexRCinflmVRfit<-IndexRCinflmVR$fit
#Extracting residuals
IndexRCinflmVRres<-IndexRCinflmVR$res
#Setting sex colorscheme
cpsex=c("darkseagreen4","lightsalmon3")
#Plotting our relationship
ggplot(IndexRCinflmVRfit, aes(x=ScIndexWorm, y=visregFit, group=recipsex, fill=recipsex, color=recipsex))+
   geom_smooth(method="loess")+
  geom_ribbon(aes(ymin=visregLwr, ymax=visregUpr), alpha=0.2)+
  geom_jitter(data=IndContactsIndexinfrecip, aes(x=ScIndexWorm, y=RecipCRs))+
  theme_classic()+
  xlab("Scaled infection intensity on index")+
  ylab("Initiated contact rate by non-index")+
  theme(text=element_text(size=19))+
  theme(legend.position="none")+
  scale_fill_manual(values=cpsex)+
  scale_color_manual(values=cpsex)+
  ylim(0, 0.07)

```

```{r visualing the infection intensity and contact rate variation relationship for the total contact glmmTMB model}
#Extracting the fit and residuals from our model 
InfIndexCR<-visreg(IndexRCinflmrecip, "ScContactInitRs", "ScIndexWorm", scale="response", partial=T, overlay=TRUE, gg=TRUE, plot=FALSE)
#Extracting the fit of our model
InfIndexCRfit<-InfIndexCR$fit
#Extracting the residuals from the model
InfIndexCRres<-InfIndexCR$res

#Plotting the relationship between received contact rate, initiated contact rate, and infection intensity 
ggplot(InfIndexCRfit, aes(x=ScContactInitRs, y=visregFit, group=ScIndexWorm, colour=ScIndexWorm, fill=ScIndexWorm))+
  geom_smooth(data=InfIndexCRfit, aes(x=ScContactInitRs, y=visregFit))+
  geom_ribbon(aes(ymin=visregLwr, ymax=visregUpr), alpha=0.2)+
  theme_classic()+
  xlab("Scaled log initiated contact rate by index")+
  ylab("Initiated contact rate by non-index")+
  theme(text=element_text(size=22))+
  theme(legend.position="none")

```

\subsubsection{Post hoc analysis for the popularity of index based on sex and index worms}

```{r Post hoc analysis for the popularity of index based on sex and index worms}
#Subsetting down to females
IndContactsIndexinfrecipF<-IndContactsIndexinfrecip%>%
  filter(recipsex == "F")

#Subsetting down to males
IndContactsIndexinfrecipM<-IndContactsIndexinfrecip%>%
  filter(recipsex == "M")

#Fitting a GLMM for females only for post hoc analysis
IndexRCinflmrecipF<-glmmTMB(RecipCRs~ScIndexWorm+
                             I(ScIndexWorm^2)+
                             ScContactInitRs+
                             I(ScContactInitRs^2)+
                             ScIndexWorm:ScContactInitRs,
                           family=beta_family(), 
                           data=IndContactsIndexinfrecipF)
#Summary for the female model model
summary(IndexRCinflmrecipF)
#Likelihood ratio test for significance in the female model
drop1(IndexRCinflmrecipF, test="Chisq")

#Fitting a GLMM for male post hoc analysis
IndexRCinflmrecipM<-glmmTMB(RecipCRs~ScIndexWorm+
                             I(ScIndexWorm^2)+
                             ScContactInitRs+
                             I(ScContactInitRs^2)+
                             ScIndexWorm:ScContactInitRs,, 
                           family=beta_family(), 
                           data=IndContactsIndexinfrecipM)
#Summary for the male model
summary(IndexRCinflmrecipM)
#Likelihood ratio test for significance in the male
drop1(IndexRCinflmrecipM, test="Chisq")

```


\subsection{Do groups with infected versus uninfected index fish change their contact rate different across days 2 and 3 and does this vary by sex?}

Subsetting down to the data we need for this analysis

```{r subsetting down for our changes in behavior analysis}
IndContacts4ch<-IndContacts4%>%
  distinct(fishID2, recipID2, .keep_all=TRUE)
#Subsetting down to females
IndContacts4chF<-IndContacts4ch%>%
  filter(Sex == "F")
#Subsetting down to males
IndContacts4chM<-IndContacts4ch%>%
  filter(Sex == "M")
```


We are fitting a linear mixed model with change in contact rate between dyas 2 and 3 for all uninfected fish in the experiment.

\subsubsection{Fitting our linear model for the change in behavior across days 2 and 3}
```{r fitting a glmm to test for differences in sex and infeciton treatment in interactions}
#fitting glmmtmb for total contacts over sex and infection treatment
Contactsmodel <- lmer(ChContactinit~
                        InfectionTrt+
                        Sex+
                        recipsex+
                        (1|population)+(1|fishID2), 
                      data=IndContacts4ch)


#Summary for our model
summary(Contactsmodel)
```


```{r validating the model fit}

sim_residuals_Contactsmodel <-simulateResiduals(Contactsmodel, 1000)  #Quantile residuals
#Ploting the quantile residuals
plot(sim_residuals_Contactsmodel) 
#Testing dispersion
testDispersion(sim_residuals_Contactsmodel)


check_model(Contactsmodel)




```

```{r testing significance of fixed effects}

Anova(Contactsmodel, type="2", test="F")

#Anova(ContactsmodelF, type="2", test="F")
#Anova(ContactsmodelM, type="2", test="F")


```


```{r ploting the relatinoships - sex, recipsex}
#Setting a labeller to change axis names in the plot
 labellerI<-labeller(InfectionTrt = c( "1" = "Infected Treatment", "0" = "Uninfected Treatment"))
#Setting colorscheme for infection and sex categorical variables
cpinf=c("darkgray", "purple")
cpsex=c("darkseagreen4","lightsalmon3")
#Exporting the fit and residual for sex in our model
ContactsmodelVR<-visreg(Contactsmodel,"Sex", scale="response", partial=T, plot=FALSE)
#Extracting fit
ContactsmodelVRfit<-ContactsmodelVR$fit
#Extracting residuals
ContactsmodelVRres<-ContactsmodelVR$res
#Plotting out the relationship
ggplot(ContactsmodelVRres, aes(x=Sex, y=visregRes, color=Sex))+
  geom_boxplot(aes(color=Sex),outliers=FALSE)+
  geom_jitter(alpha=.5, size=.5)+
  theme_classic()+
  theme(text=element_text(size=22))+
  ylab("Change in contact initiated")+
  xlab("Sex of the initiator")+
  scale_color_manual(values=cpsex)+
  theme(legend.position="none")


#Extracting residuals and fit for the recipient sex in the model
ContactsmodelrecipVR<-visreg(Contactsmodel, "recipsex", scale="response", partial=T, plot=FALSE)
#Extracting fit
ContactsmodelrecipVRfit<-ContactsmodelrecipVR$fit
#Extracting residuals
ContactsmodelrecipVRres<-ContactsmodelrecipVR$res
#Plotting the residuals
ggplot(ContactsmodelrecipVRres, aes(x=recipsex, y=visregRes))+
  geom_boxplot(aes(color=recipsex),outliers=FALSE)+
  geom_jitter(alpha=.5, size=.5)+
  theme_classic()+
  theme(text=element_text(size=22))+
  ylab("Change in contact initiated")+
  xlab("Sex of the recipient")+
  scale_color_manual(values=cpsex)+
  theme(legend.position="none")
```

\subsubsection{Post-hoc analyses for exploring sex differences in the change in contact rate between days 2 and 3}


```{r posthoc analysis for change in contact rate}
#Fitting our post-hoc analysis for female change in contact model
ContactsmodelF <- lmer(ChContactinit~
                        (1|population)+(1|fishID2), 
                      data=IndContacts4chF)
#Summary for the female model
summary(ContactsmodelF)

#One sample t-test to test if the change in contact rate is different than zero
t.test(IndContacts4chF$ChContactinit)

#FItting our post-hoc analysis for male change in contact model
ContactsmodelM <- lmer(ChContactinit~
                        (1|population)+
                        (1|fishID2), 
                      data=IndContacts4chM)

#Summary for the male model
summary(ContactsmodelM)

#One sample t-test to test if the change in contact rate is different than zero
t.test(IndContacts4chM$ChContactinit)
```

